Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-20 prop_iop.tex
prop_letter.tex \section{Gleason} Updating...
prop_letter.tex Updating comments...

\begin{proof}Given two events~$E_{0},E_{1}\subseteq\Omega$, we
want to verify 
\begin{equation}
\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\,.\label{eq:classical-convex}
\end{equation}
When $E_{0}\cap E_{1}=\emptyset$, equation~(\ref{eq:classical-convex})
is a special case of equation~(\ref{eq:classical-include}). When
$E_{0}\cup E_{1}=\Omega$, we have $E_{0}^{c}\cap E_{1}^{c}=\emptyset$,
where $E_{0}^{c}=\Omega\backslash E_{0}$ and $E_{1}^{c}=\Omega\backslash E_{1}$.
Thus, 
\begin{equation}\eqalign{ 
& \bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\ 
\Rightarrow & \bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cup E_{1}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cap E_{1}^{c}\right)\right)\subseteq\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{1}^{c}\right)\right)\\ 
\Rightarrow & \bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\,. 
}\end{equation}
When the number of elements in $\Omega\le3$, either $E_{0}\cap E_{1}=\emptyset$
or $E_{0}\cup E_{1}=\Omega$ must happen so that $\bar{\mu}^{\Omega}$
is convex. \end{proof}

@@

\subsection{Expectation Values for $0$-deterministic CIVPMs}

Finally, we are going to focus on CIVPMs whose range is $\mathscr{I}_{\mathrm{D}}$.

\begin{lemma}\label{lemma:deterministic-CIVPM}Given a CIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$,
the following statements are true:
\begin{enumerate}
\item Its left-end~$\mul$ and right-end~$\mur$ are the same CRVPM, and
will be simply denoted by $\pmeas:\events\rightarrow\left\{ 1,0\right\} $.
\item There is a unique $\omega_{0}\in\Omega$ such that $\bar{\mu}\left(\left\{ \omega_{0}\right\} \right)=\left\{ \pmeas\left(\left\{ \omega_{0}\right\} \right)\right\} =\necess$,
denoted by $\pmeas^{-1}\left(1\right)$ or $\bar{\mu}^{-1}\left(\necess\right)$.
\end{enumerate}
\end{lemma}

\begin{proof}~
\begin{enumerate}
\item The directly consequence of lemma~\ref{lemma:QIVPM2QCM}.
\item Since the range of $\pmeas$ is $\left\{ 1,0\right\} $ and $\sum_{\omega\in\Omega}\pmeas\left(\left\{ \omega\right\} \right)=1$,
there is a unique $\omega_{0}$ such that $\mu\left(\left\{ \omega_{0}\right\} \right)=1$
which implies $\bar{\mu}\left(\left\{ \omega_{0}\right\} \right)=\necess$.
\end{enumerate}
\end{proof}

>>>>

\begin{proof}Given two events~$E_{0},E_{1}\subseteq\Omega$, we
want to verify 
\begin{equation}
\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\,.\label{eq:classical-convex}
\end{equation}
When $E_{0}\cap E_{1}=\emptyset$, equation~(\ref{eq:classical-convex})
is a special case of equation~(\ref{eq:classical-include}). When
$E_{0}\cup E_{1}=\Omega$, we have $E_{0}^{c}\cap E_{1}^{c}=\emptyset$,
where $E_{0}^{c}=\Omega\backslash E_{0}$ and $E_{1}^{c}=\Omega\backslash E_{1}$.
Thus, 
\begin{equation}\eqalign{ 
& \bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\ 
\Rightarrow & \bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cup E_{1}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cap E_{1}^{c}\right)\right)\subseteq\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{1}^{c}\right)\right)\\ 
\Rightarrow & \bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\,. 
}\end{equation}
When the number of elements in $\Omega\le3$, one of the following
must happen
\begin{enumerate}
\item $E_{0}\cap E_{1}=\emptyset$.
\item $E_{0}\cup E_{1}=\Omega$.
\item $E_{0}\subseteq E_{1}$ or $E_{1}\subseteq E_{0}$.
\end{enumerate}
so that $\bar{\mu}^{\Omega}$ is convex. \end{proof}

@@

\subsection{Expectation Values for $\delta$-deterministic CIVPMs}

Finally, we are going to focus on $\delta$-deterministic CIVPMs.

\begin{lemma}\label{lemma:deterministic-CIVPM}Consider a sample
space~$\Omega$ with $d$ elements, and $\delta<1/d$. Given a $\delta$-deterministic
CIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, the following statements
are true:
\begin{enumerate}
\item If $\bar{\mu}$ is $0$-deterministic, its left-end~$\mul$ and right-end~$\mur$
are the same CRVPM, and will be simply denoted by $\pmeas:\events\rightarrow\left\{ 1,0\right\} $.
\item There is a unique element~$\omega_{0}\in\Omega$ such that $\bar{\mu}\left(\left\{ \omega_{0}\right\} \right)\subseteq\left[1-\delta,1\right]$,
denoted by $\bar{\mu}^{-1}\left(\left[1-\delta,1\right]\right)$.
In particular, if $\bar{\mu}$ is $0$-deterministic, $\bar{\mu}\left(\left\{ \omega_{0}\right\} \right)=\left\{ \pmeas\left(\left\{ \omega_{0}\right\} \right)\right\} =\necess$,
denoted by $\pmeas^{-1}\left(1\right)$ or $\bar{\mu}^{-1}\left(\necess\right)$.
\end{enumerate}
\end{lemma}

\begin{proof}~
\begin{enumerate}
\item The directly consequence of lemma~\ref{lemma:QIVPM2QCM}.
\item We have either $\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[0,\delta\right]$
or $\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[1-\delta,1\right]$.
\begin{casenv}
\item If $\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[0,\delta\right]$
for all $\omega\in\Omega$, then $\necess=\bar{\mu}\left(\Omega\right)\subseteq\sum_{\omega\in\Omega}\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[0,d\delta\right]\Rightarrow d\delta>1$.
A contradiction!
\item If there are more than one~$\omega\in\Omega$ such that $\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[1-\delta,1\right]$,
then $\necess=\bar{\mu}\left(\Omega\right)\subseteq\sum_{\omega\in\Omega}\bar{\mu}\left(\left\{ \omega\right\} \right)\subseteq\left[2-2\delta,d\right]\Rightarrow2\delta>1$.
A contradiction!
\end{casenv}
Therefore, there is a unique element~$\omega_{0}\in\Omega$ such
that $\bar{\mu}\left(\left\{ \omega_{0}\right\} \right)\subseteq\left[1-\delta,1\right]$.
\end{enumerate}
\end{proof}

### 2017-09-20 prop_letter.tex
prop_letter.tex Updating comments...

A principal result is that recasting the Kochen-Specker theorem in
terms of ``$\delta$-determinism'', with~$\delta$ quantifying the
effect of finite-precision uncertainties on measurement outcomes,
reducing to the purely deterministic Kochen-Specker case for
$\delta=0$, can resolve the question of the effect of finite precision
and imperfections on contextuality.  A remarkable consequence of our
analysis is that not only does the quantum-mechanical requirement for
contextuality continue to hold, thus maintaining the Kochen-Specker
result for a \emph{non-vanishing} range of $\delta$, but that at a
\emph{certain fixed value} of $\delta$, there is a sharp transition,
parallel in its consequences to, say, a phase transition, to
\emph{non-contextuality}; that is, suddenly a quantum system can be
simulated by a non-contextual hidden variable theory, turning the
system effectively from a quantum system into a classical system. 
% that is, quantum mechanics suddenly fails to hold, and all variables
% can be specified simultaneously, turning the theory effectively into a
% classical theory that can be specified consistently by hidden
% variables.
% \yutsung{The part after ``that is''
%   looks strange.
% \begin{enumerate}
% \item The Kochen-Specker theorem is not about quantum mechanics, is about whether
% quantum mechanics can be simulated by a non-contextual deterministic hidden
% variable theory or not. Hence, ``quantum mechanics suddenly fails to hold''
% doesn't make a lot sense.
% \item The question is whether there exist a QIVPM means ``a classical theory
% consistently by hidden variables.'' At least, a QIVPM is not deterministic, and
% I need to check whether it qualities as a classical probabilistic hidden variable
% theory\ldots(?)
% \end{enumerate}
% }
\yutsung{Still not comfortable about the current statement...}
These results provide a new insight into the Meyer-Mermin debate by
presenting a theory in which there is a \emph{parameter} interpolating
between what appeared previously to be irreconcilable interpretations
of contextuality.

@@

\yutsung{About contextuality or non-contextuality, summarize the
existence and non-existence of $\delta$-deterministic QIVPMs in Fig.~\ref{fig:contextuality-non-contextuality}.}
\begin{figure*}
\[
\xymatrix{ & \delta=0 &  & \delta=\frac{1}{3} & \delta=1\\
\textrm{Classical IVPMs} & \ar[rrr] &  &  & \,\\
\textrm{Instance} & \bar{\mu}_{4} &  & \bar{\mu}_{5},\bar{\mu}_{6} & \bar{\mu}_{2},\textrm{ every classical IVPM}\\
\textrm{Interpertation} & \textrm{D Classical Theory} &  & \textrm{P Classical Theory?} & \textrm{P Classical Theory?}\\
\textrm{QIVPMs} & \ar@2{|-|}[rr]_{\textrm{NE by Thm. \ref{cor:Kochen-Specker-IVPM}}}\ar[rrr] &  &  & \,\\
\textrm{Instance} & \textrm{NE by Thm. }\ref{thm:Kochen-Specker} &  & \iota_{\infty}\circ\mu_{\mathbb{1}/3}^{\mathrm{B}} & \bar{\mu}_{1},\bar{\mu}_{3}\textrm{, every QIVPM}\\
\textrm{Interpertation} & \textrm{NEDNCHVT}\ar@{}[rr]|-{\textrm{NEPNCHVT?}} &  & \textrm{EPNCHVT?} & \textrm{EPNCHVT?}
}
\]
\caption{\label{fig:contextuality-non-contextuality}The figure summarize the
existence and non-existence of $\delta$-deterministic classical IVPMs
and QIVPMs. In the figure, ``D'' denotes ``deterministic,'' P
denotes ``probabilitic,'' ``(N)E'' denotes ``(Non-)Exist,''
``(N)C'' denotes ``(Non-)Contextuality,'' and HVT denotes ``Hidden-Variable
Theory.''}
\end{figure*}

>>>>

A principal result is that recasting the Kochen-Specker theorem in
terms of ``$\delta$-determinism'', with~$\delta$ quantifying the
effect of finite-precision uncertainties on measurement outcomes,
reducing to the purely deterministic Kochen-Specker case for
$\delta=0$, can resolve the question of the effect of finite precision
and imperfections on contextuality.  A remarkable consequence of our
analysis is that not only does the quantum-mechanical requirement for
contextuality continue to hold, thus maintaining the Kochen-Specker
result for a \emph{non-vanishing} range of $\delta$, but that at a
\emph{certain fixed value} of $\delta$, there is a sharp transition,
parallel in its consequences to, say, a phase transition, to
\emph{non-contextuality}; that is, suddenly a quantum system can be
simulated by a non-contextual hidden variable theory, turning the
system effectively from a quantum system into a classical system. 
% that is, quantum mechanics suddenly fails to hold, and all variables
% can be specified simultaneously, turning the theory effectively into a
% classical theory that can be specified consistently by hidden
% variables.
% \yutsung{The part after ``that is''
%   looks strange.
% \begin{enumerate}
% \item The Kochen-Specker theorem is not about quantum mechanics, is about whether
% quantum mechanics can be simulated by a non-contextual deterministic hidden
% variable theory or not. Hence, ``quantum mechanics suddenly fails to hold''
% doesn't make a lot sense.
% \item The question is whether there exist a QIVPM means ``a classical theory
% consistently by hidden variables.'' At least, a QIVPM is not deterministic, and
% I need to check whether it qualities as a classical probabilistic hidden variable
% theory\ldots(?)
% \end{enumerate}
% }
\yutsung{Still not comfortable about the current statement... Notice
that the key properties of the non-contextual hidden variable theory
used in our Mermin-Peres square proof is
\begin{itemize}
\item The expectation value of an observable is one of the eigenvalue of
that observable.
\item For $0$-deterministic QIVPMs, we have
\begin{equation}
\prod_{j}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j}\mathbf{O}_{j}}_{\bar{\mu}}\,.\label{eq:Kochen-Specker-Rules}
\end{equation}
\end{itemize}
I think these two properties might be weakened when $0<\delta<\frac{1}{3}$,
and completely break down when $\delta\ge\frac{1}{3}$. (Need a proof!)
If this is the case, the correct statement is that ``the assumption
of non-contextuality hidden variable theory could be considered hold
when $\delta<\frac{1}{3}$, and is completely void when $\delta\ge\frac{1}{3}$.''
So we should not even ask ``whether the existence of QIVPMs when
$\delta\ge\frac{1}{3}$ implies non-contextuality hidden variable
theory or not.'' A little bit different analog could be: we should
not ask ``where is the momentum of a particle when we measure its
position...''}

These results provide a new insight into the Meyer-Mermin debate by
presenting a theory in which there is a \emph{parameter} interpolating
between what appeared previously to be irreconcilable interpretations
of contextuality.

@@
